Flute and kink instabilities in a dynamically twisted flux tube with anisotropic plasma viscosity

### 2022-01-11 paper.tex
?
few minor edits, mainly testing my git

\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc} % modern encoding

\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
%\usepackage{amssymb}	% Extra maths symbols

@@

\newcommand{\rev}{{\color{red} #1}}

@@

The helical kink instability is a form of ideal magnetohydrodynamic (MHD)
instability which occurs in highly twisted magnetic flux tubes such as those
making up much of the solar corona~\citep{realeCoronalLoopsObservations2014}
and has been well studied in the coronal
context~\citep{hoodKinkInstabilitySolar1979,hoodCoronalHeatingMagnetic2009,browningSolarCoronalHeating2003b,barefordShockHeatingNumerical2015,quinnEffectAnisotropicViscosity2020}.
Given its energetic nonlinear development, it is considered a potential
mechanism for heating the solar corona through the theory of
nanoflares~\citep{klimchukSolvingCoronalHeating2006,browningMechanismsSolarCoronal1991}
and a key mechanism in the production of solar
flares~\citep{hoodKinkInstabilitySolar1979}. \rev{Some of our previous work} has
investigated a twisted magnetic flux tube already linearly unstable to the
helical kink instability, focussing specifically on the effect of anisotropic
viscosity on the nonlinear
dynamics~\citep{quinnEffectAnisotropicViscosity2020}.
In~\citet{quinnEffectAnisotropicViscosity2020} and most other investigations of
the kink instability e.g.~\citet{hoodCoronalHeatingMagnetic2009},
a perturbation is applied to an already significantly twisted flux tube. An
alternative way to excite the instability (and the way employed here) is to
start with an initially straight field and apply twisting motions at the
boundaries to form a twisted flux tube which eventually becomes unstable. This
kind of dynamic excitation of the kink instability is useful in that it
represents more closely actual evolution of magnetic flux tubes and the
associated instabilities in the solar corona. In our simulations, the dynamic
twisting of the flux tube reveals an additional instability, the flute
instability, which has been theorised (for example,
in~\citet{priestMagnetohydrodynamicsSun2013}). While oscillations resembling
flute perturbations have been found in simulations of coronal
loops~\citep{terradasEffectMagneticTwist2018}, to our knowledge, this is the
first time the flute instability has been investigated computationally in
a coronal context.

>>>>

%\usepackage{newtxtext,newtxmath}
%\usepackage[T1]{fontenc} % modern encoding

\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

@@

\newcommand{\rev}[1]{{\color{red} #1}}

@@

The helical kink instability is a form of ideal magnetohydrodynamic (MHD)
instability which occurs in highly twisted magnetic flux tubes such as those
making up much of the solar corona~\citep{realeCoronalLoopsObservations2014}
and has been well studied in the coronal
context~\citep{hoodKinkInstabilitySolar1979,hoodCoronalHeatingMagnetic2009,browningSolarCoronalHeating2003b,barefordShockHeatingNumerical2015,quinnEffectAnisotropicViscosity2020}.
Given its energetic nonlinear development, it is considered a potential
mechanism for heating the solar corona through the theory of
nanoflares~\citep{klimchukSolvingCoronalHeating2006,browningMechanismsSolarCoronal1991}
and a key mechanism in the production of solar
flares~\citep{hoodKinkInstabilitySolar1979}. \rev{In our previous work
we have} investigated a twisted magnetic flux tube already linearly unstable to the
helical kink instability, focussing specifically on the effect of anisotropic
viscosity on the nonlinear
dynamics~\citep{quinnEffectAnisotropicViscosity2020}.
\rev{There and in} most other investigations of
the kink instability e.g.~\rev{\citep{hoodCoronalHeatingMagnetic2009}},
a perturbation is applied to an already significantly twisted flux tube. An
alternative way to excite the instability (and the way employed here) is to
start with an initially straight field and apply twisting motions at the
boundaries to form a twisted flux tube which eventually becomes unstable. This
kind of dynamic excitation of the kink instability is useful in that it
represents more closely the actual evolution of magnetic flux tubes and the
associated instabilities in the solar corona. In our simulations, the dynamic
twisting of the flux tube reveals an additional instability, the flute
instability, which has been theorised, for example
in~\rev{\citep{priestMagnetohydrodynamicsSun2013}}. While oscillations resembling
flute perturbations have been found in simulations of coronal
loops~\citep{terradasEffectMagneticTwist2018}, to our knowledge, this is the
first time the flute instability has been investigated computationally in
a coronal context.
